---
title: "Oct6"
output: pdf_document
date: "2022-11-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library("tidyverse")
library("dplyr")
source("../constant/constant.R")
source("../function/disc_gamma.R")
source("../function/make_plot.R")
source("../model/pls/ridge.R")


### Importing Rt estimation package
# library(EpiEstim)
source("../run/run_EpiEstim.R")

source("../model/pls/ridge.R")

source("../run/run_pf_all.R")

source("../run/run_ridge_gd.R")

### Importing data, no cyclic effect

# d = read.csv("../data/processed/d2.csv")
# plot(d$y, type = "l")
# plot(d$r, type = "l")

```
### Run methods

## EpiEstim
epiestim <- run_epiestim(d, p1 = sid_covid_mean, p2 = sid_covid_sd)
plot(epiestim, type = "l")

## PLS

# CLosed form
lamb <- cv_loss(d$iwt, d$y)
# loocv <- CV(d$iwt, d$y, lambdas = exp(seq(from=0.1, to = 10, by = 0.3)))
ridge <- get_r(d$iwt, d$y, lambda = lamb)
plot(ridge, type = "l")

# NLP GD
ridge_gd <- run_ridge_gd(d, lamb)
plot(ridge_gd$estimate, type = "l")

## Particle Filter
pf <- run_pf(d, 0.5)
plot(pf$filter.mean.x, type = "l")

## EpiNow
epin <- read.csv("data/results/epinow2_d2.csv")
plot(epin$median[1:500], type = "l")



### Build Rt estimation dataframe

all_rt <- data.frame(idx = 1:nrow(d), a_True_Rt = d$r, EpiEstim=epiestim, Ridge_closed_form = ridge, Ridge_gd = ridge_gd$estimate, Particle_Filter = pf$filter.mean.x, EpiNow2=epin$median[1:500])

all_rt %>%
  pivot_longer(!idx, names_to = "method", values_to = "vals")%>%
  ggplot(aes(x = idx, fill=method, linetype=method, color = method))+
  geom_line(aes(y=vals))+
  theme_bw()+
  xlab("Time")+
  ylab("Rt")


```