---
title: "Oct19"
output: pdf_document
date: '2022-10-19'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PLS closed-form solution




\noindent 1. Deriving the closed form solution for PLS with MSE loss and L2 penalty on consecutive smoothing $R_t$. Starting from the objectives in the matrix form.\


$\Vert I-RW\Vert^2_2 + \lambda \Vert R_{t+1} - R_t \Vert^2_2$

$=> \Vert I-RW\Vert^2_2 + \lambda \Vert DR \Vert^2_2$

where $D \in R^{(d-1)*d}$ is the difference matrix, with all 1's on the diagonal, and -1 on the off-diagonal above. 

where $W$ is a square matrix with diagonol being the vector $w_t = \sum_{a=1}^t I_{t-a}s_a$

$=>W^TR^TRW -2W^TR^TI + I^TI + \lambda R^TD^TDR$\

Taking derivative and equate to 0

$=> 2W^TR^TW - 2W^TI + 2\lambda R^TD^TD = 0$

$=> R^T(W^TW + \lambda D^TD) = W^TI$

$=> R^T = (W^TW + \lambda D^TD)^{-1}W^TI$


```{r, fig.show="hold", out.width="50%"}
### Import s1
s1 <- read.csv("../data/processed/a.csv")

### Plotting
plot(s1$r)
plot(s1$y)
```

```{r, output=FALSE}
r1_length <- nrow(s1)

### Construct difference matrix
D = diag(r1_length)
D[row(D) == col(D)-1] = -1
D = D[1:(nrow(D)-1),]


### Construct the Iwt vector
source("../function/get_iwt.R")
source("../function/disc_gamma.R")
source("../constant/constant.R")

s1_iwt <- get_iwt(s1$y, disc_gamma(1:nrow(s1), sid_ebola_shape, sid_ebola_scale))
W <- diag(s1_iwt)
Y <- s1$y

# R = solve((t(W)%*%W-lambda*t(D)%*%D))%*%t(W)%*%Y

get_r = function(Y,W,D, lambda=10) solve((t(W)%*%W-lambda*t(D)%*%D))%*%t(W)%*%Y

get_obj = function(Y,W,R,D) sum(Y-W%*%R)^2 + sum(lambda*(D%*%R)^2) 

```


```{r}

### Function: Cross validation on ridge regression
CV <- function(W, Y, D, lambdas = exp(seq(0.1,10,0.2))){
  ### Length of data
  dat_length = length(Y)
  
  ### Get grid of lambda
  cv_scores = c()
  best = 0
  
  I = diag(rep(1, dat_length))
  
  for (lambda in lambdas){
    L = solve((t(W)%*%W-lambda*t(D)%*%D))%*%t(W)
    H = I-t(L)%*%W
    H_tilde = diag(diag(H))
    HHY = H%*%solve(H_tilde)%*%Y
    E_cv = 1/(dat_length)*sum(HHY^2)
    cv_scores = c(cv_scores, E_cv)
  }
  
  return(cv_scores)
  
}

lambdas = exp(seq(0.1,10,0.2))

cv_scores = CV(W, Y, D, lambdas = lambdas)
plot(log(cv_scores))
```

```{r, fig.show="hold", out.width="50%"}
ordered = order(cv_scores)
lambdas = lambdas[ordered]

### Plot the best 5
for (optim_lambda in lambdas[1:5]){
  R = get_r(Y, W, D, optim_lambda)
  plot(R, type="l")
}


```


```{r}
source("../function/process_data.R")
library("zoo")
ca <- get_owid_data(data_loc = "../data/raw/owid_Sep5.csv")
ca$y[which.max(ca$y)] = NA
ca$y <- na.locf(ca$y)
ca <- ca[100:(nrow(ca)-100),]
plot(ca$y)
```

```{r, fig.show="hold", out.width="50%"}
ca_iwt <- get_iwt(ca$y, disc_gamma(1:nrow(ca), sid_covid_shape, sid_covid_scale))
# ca_iwt[1] <- ca_iwt[2]
ca_iwt <- na.locf(ca_iwt)

length(ca_iwt)

W <- diag(ca_iwt)
Y <- ca$y
ca_length <- nrow(ca)

D = diag(ca_length)
D[row(D) == col(D)-1] = -1
D = D[1:(nrow(D)-1),]

print(dim(W))
print(length(Y))
print(dim(D))

R_ca <- get_r(Y, W, D, lambda=10000)
plot(R_ca, type="l")

lambdas = exp(seq(2,10,0.2))
CV_ca <- CV(W, Y, D, lambdas)


### Plot the best 5
for (optim_lambda in lambdas[1:5]){
  optim_lambda
  R = get_r(Y, W, D, optim_lambda)
  plot(R, type="l")
}

```

```{r}
source("../function/process_data.R")
source("../function/gen_data.R")

mp <- get_owid_mp(data_loc = "../data/raw/owid-monkeypox-data.csv")
head(mp)
plot(mp$y, type="l")
mp_length <- nrow(mp)

sid_mp_mean
sid_mp_sd

sid_mp_shape = gamma_reparam(sid_mp_mean, sid_mp_sd)[1]
sid_mp_scale = gamma_reparam(sid_mp_mean, sid_mp_sd)[2]

sid_mp_shape
sid_mp_scale
mp_iwt <- get_iwt(mp$y, disc_gamma(1:nrow(mp), sid_mp_shape, sid_mp_scale))


D = diag(mp_length)
D[row(D) == col(D)-1] = -1
D = D[1:(nrow(D)-1),]

Y = mp$y

W = diag(mp_iwt)


CV_mp = CV(W, Y, D)

plot(log(CV_mp))

lambdas = exp(seq(0.1,13,0.2))

ordered = order(CV_mp)
lambdas = lambdas[ordered]

for (optim_lambda in lambdas[1:5]){
  R = get_r(Y, W, D, optim_lambda)
  plot(R, type="l")
}
# 
# R_mp <- get_r(Y, W, D, lambda=50)
# 
# plot(R_mp, type="l")

```







# Smoothness at head and tail

1. Start and end of the synthetic datasets have low case counts, therefore prediction (using renewal equation) fluctuate. Model should take this into consideration and should not put too much effort in trying to smoothing those regions.

2. Penalty change with size of case count?

3. Adding cyclic penalty term?







