---
title: "Oct27"
output:
  pdf_document: default
  html_document: default
date: '2022-10-27'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, fig.show="hold", out.width="50%"}

### Importing dataset ###
source("../function/get_iwt.R")
source("../function/disc_gamma.R")
source("../constant/constant.R")

s5 <- read.csv("../data/processed/d1.csv")
# s5_iwt <- get_iwt(s5$y, disc_gamma(1:nrow(s5), sid_ebola_shape, sid_ebola_scale))
s5_iwt <- s5$iwt

plot(s5$r)
plot(s5$y, type = "l")
```


# Previous work


Last time, I let the penalty to change as a function of the incidence cases for the following reasons:

1. (Primary) $R_t$ estimation in lower case count areas (beginning and ending) fluctuates a lot. This is expected, but it forces a uniformly large penalty on the whole estimation. 

2. For the synthetic dataset, I let the amplitude of the cycle to change as a function of the incidence case counts.


The smoothing function from last time is $F* (DR) \odot(DR)$, where $\odot$ is the hadamard product (element-wise product), where $F = f(W)$. Example used last time has $F = log(w+1)$

```{r}
r_smooth_c1 <- function(r, iwt, pen_func){
  return(sum((pen_func(iwt[1:length(iwt)-1]+1)*diff(r))^2))
}
```



The ridge objective function is
```{r}

ridge_obj <- function(data, par, loss_func, iwt = iwt, smooth_func, penalties, pen_func = log){
  
  dat_length = nrow(data)
  
  loss = loss_func(z=data$y, iwt = iwt, r = par)
  
  r_pen <- penalties$r* smooth_func(par, iwt, pen_func)
  
  obj_value = sum(loss+r_pen)
  
  return(obj_value)
}

```


```{r}
init_r = rep(1.5, nrow(s5))
```

```{r}
source("../model/pls/penalties_smooth.R")

result <- nlm(f=ridge_obj, p = init_r, iterlim =2000, print.level = 0, data=s5, penalties = list("r"=1), loss_func = rmse_loss, smooth_func = r_smooth_c1, pen_func = sqrt, iwt = s5_iwt)
```


The result looks like

```{r, fig.show="hold", out.width="50%"}
source("../function/make_plot.R")
diags <- diag_plots(s5$r, result$estimate, s5_iwt, s5$y, cap=0)
diags$rt
diags$oneday
```

Comparing to the closed form solution:

```{r, fig.show="hold", out.width="50%"}
# source("../model/pls/ridge.R")
# source("../function/make_plot.R")
# cv <- CV(s5_iwt, s5$y)
# plot(log(cv$scores))
# ridge_r <- get_r(s5_iwt, s5$y, lambda=cv$lambdas[1])
# 
# cv$lambdas[1]
# diags_ridge <- diag_plots(s5$r, ridge_r, s5_iwt, s5$y, cap=100)
# diags_ridge$rt
# diags_ridge$oneday
```



